A New Method of Modelling Tuneable Lasers with Functional Composition

### 2019-05-29 Thesis/Appendices/Asymptotics.tex
Finished second draft

We shall now consider the other limit, when $s \rightarrow \infty$. Using a similar substitution we instead write \eqref{eq:var} as
\begin{align}
\label{eq:asymp2}
\frac{1}{4} \eps x^4 = -x^3 + 5x^2 - 8x + 4,
\end{align}
where $x = \sigma^2$, and $\eps = s^{-4}$ instead so that we still have $\eps \rightarrow 0$. Furthermore, we must make a correction to the series expansion:
\begin{align}
\label{eq:series2}
x = \eps^{-\xi} \left( x_0 + x_1 \eps^\alpha + x_2 \eps^\beta + \cdots \right).
\end{align}
The reason for this is because the equation is now singular---when $\eps=0$ the equation transforms from a quartic to a cubic, losing a root. As with before, we obtain dominant balance when the quartic term balances with the cubic term which occurs when $1 - 4 \xi = -3 \xi$, or $\xi = 1$. Then to $\bigO{\eps^{-3}}$, expression \eqref{eq:asymp2} gives
\begin{align*}
x_0^3 (x_0 + 4) &= 0,
\end{align*}
we shall choose $x_0 = 0$---since $x \geq 0$. Continuing, we obtain a dominant balance when $\alpha = 1$. So the next condition occurs at $\bigO{1}$ where
\begin{align*}
0 = -x_1^3 + 5x_1^2 - 8x_1 + 4 = -(x_1 - 1)(x_1 - 2)^2,
\end{align*}
so either $x_1 = 1$ or $x_1 = 2$, we shall see in the next step that we can eliminate one of these. The next term obtains dominant balance when the left hand side balances with the linear term of the right hand side, that is, when $\beta = 2$. Now at $\bigO{\eps}$,
\begin{align*}
\frac{1}{4} x_1^4 = -3x_1^2 x_2 + 5\cdot2 x_1 x_2 - 8x_2 = -x_2 (3x_1 - 4)(x_1 - 2),
\end{align*}
where again, the coefficients come from the binomial expansions. From this it is clear that $x_2$ is unique provided $x_1 \ne 2$ and
\begin{align*}
x_2 &= \frac{-x_1^4}{4(3x_1 - 4)(x_1 - 2)}.
\end{align*}
Thus, $x_1 = 1$, $x_2 = -\frac{1}{4}$, and
\begin{align*}
x = 1 - \frac{1}{4} \eps + x_2 \eps^{\gamma - 1} + \cdots.
\end{align*}
Expanding one final term we find $\gamma = 3$, and to $\bigO{\eps^2}$,
\begin{align*}
\frac{1}{4} 4x_2 x_1^3 &= -(3x_3 x_1^2 + 3x_2^2 x_1) + 5(2x_3 x_1 + x_2^2) - 8x_3.
\end{align*}
Knowing the values of $x_1$, and $x_2$ this is simply an arithmetical calculation yielding $x_3 = \frac{3}{8}$. Finally,
\begin{align*}
x &= 1 - \frac{1}{4}\eps + \frac{3}{8}\eps^2 + \bigO{\eps^3},
\end{align*}
or,
\begin{align*}
\sigma^2 &= 1 - \frac{1}{4s^4} + \frac{3}{8s^8} + \bigO{s^{-12}}
\end{align*}
this approximation is shown in Figure \ref{fig:limsinfty}. \\
\begin{figure}[tbp]
\input{./Figures/Lim_s_Infty}
\caption{Asymptotic expansion of the variance as $s \rightarrow \infty$.}
\label{fig:limsinfty}
\end{figure}

In conclusion, the asymptotic expansions of the observables are
\begin{align*}
\sigma^2 &=
\begin{cases*}
2s(1 - s) + \bigO{s^3} & $s \rightarrow 0$ \\
1 - \frac{1}{4s^4} + \frac{3}{8s^8} + \bigO{s^{-12}} & $s \rightarrow \infty$
\end{cases*}
& \phi &=
\begin{cases*}
\frac{1}{2}s - \frac{1}{24}s^3 + \bigO{s^5} & $s \rightarrow 0$ \\
\frac{\pi}{4} - \frac{1}{2s^2} + \frac{5}{12s^6} + \bigO{s^{-10}} & $s \rightarrow \infty$
\end{cases*} \\
C &=
\begin{cases*}
1 - s + \frac{1}{2}s^2 + \bigO{s^3} & $s \rightarrow 0$ \\
\frac{1}{2s^2} - \frac{3}{8s^6} + \bigO{s^{-10}} & $s \rightarrow \infty$
\end{cases*}
& \zeta &=
\begin{cases*}
1 - \frac{1}{2}s + \frac{1}{8}s^2 + \bigO{s^3} & $s \rightarrow 0$ \\
\frac{1}{4^{1/4}s} - \frac{3 \cdot 4^{3/4}}{32 s^5} + \bigO{s^{-9}} & $s \rightarrow \infty$.
\end{cases*}
\end{align*}
The general form of these is as expected. When $s \rightarrow 0$, we effectively remove the CFBG from the cavity. This causes the pulse to be perpetually modulated, and so, the pulse approaches a $\delta$ function, and therefore, $\sigma^2 \rightarrow 0$. Moreover, since $\zeta$ is the ratio of the amplitude of the pulse before and after dispersion, $\zeta \rightarrow 1$. Similarly, $\phi \rightarrow 0$ since nothing in the circuit has the capacity to provide a phase shift. In the opposite extreme, $s \rightarrow \infty$, we have an incredibly long / powerful CFBG---this causes the pulse to broaden greatly, while conserving its energy. Because of this, the variance approaches unity because the shape is entirely determined by the modulation function. Additionally, $\zeta \rightarrow 0$ because the pulse becomes infinitely wide, and thus, the amplitude becomes infinitesimal.

>>>>

We shall now consider the other limit, when $s \rightarrow \infty$. Using a similar substitution we instead write \eqref{eq:var} as
\begin{align}
\label{eq:asymp2}
\frac{1}{4} \eps x^4 = -x^3 + 5x^2 - 8x + 4,
\end{align}
where $x = \sigma^2$, and $\eps = s^{-4}$ instead so we still have $\eps \rightarrow 0$. Furthermore, we must make a correction to the series expansion:
\begin{align}
\label{eq:series2}
x = \eps^{-\xi} \left( x_0 + x_1 \eps^\alpha + x_2 \eps^\beta + \cdots \right).
\end{align}
The reason for this is because the equation is now singular---when $\eps=0$ the equation transforms from a quartic to a cubic, losing a root. As with before, we obtain dominant balance when the quartic term balances with the cubic term which occurs when $1 - 4 \xi = -3 \xi$, or $\xi = 1$. Then to $\bigO{\eps^{-3}}$, expression \eqref{eq:asymp2} gives
\begin{align*}
x_0^3 (x_0 + 4) &= 0,
\end{align*}
we shall choose $x_0 = 0$---since $x \geq 0$. Continuing, we obtain a dominant balance when $\alpha = 1$. So the next condition occurs at $\bigO{1}$ where
\begin{align*}
0 = -x_1^3 + 5x_1^2 - 8x_1 + 4 = -(x_1 - 1)(x_1 - 2)^2,
\end{align*}
so either $x_1 = 1$ or $x_1 = 2$, we shall see in the next step that we can eliminate one of these. The next term obtains dominant balance when the left hand side balances with the linear term of the right hand side, that is, when $\beta = 2$. Now at $\bigO{\eps}$,
\begin{align*}
\frac{1}{4} x_1^4 = -3x_1^2 x_2 + 5\cdot2 x_1 x_2 - 8x_2 = -x_2 (3x_1 - 4)(x_1 - 2),
\end{align*}
where again, the coefficients come from the binomial expansions. From this it is clear that $x_2$ is unique provided $x_1 \ne 2$ and
\begin{align*}
x_2 &= \frac{-x_1^4}{4(3x_1 - 4)(x_1 - 2)}.
\end{align*}
Thus, $x_1 = 1$, $x_2 = -\frac{1}{4}$, and
\begin{align*}
x = 1 - \frac{1}{4} \eps + x_2 \eps^{\gamma - 1} + \cdots.
\end{align*}
Expanding one final term we find $\gamma = 3$, and to $\bigO{\eps^2}$,
\begin{align*}
\frac{1}{4} 4x_2 x_1^3 &= -(3x_3 x_1^2 + 3x_2^2 x_1) + 5(2x_3 x_1 + x_2^2) - 8x_3.
\end{align*}
Knowing the values of $x_1$, and $x_2$ this is simply an arithmetical calculation yielding $x_3 = \frac{3}{8}$. Finally,
\begin{align*}
x &= 1 - \frac{1}{4}\eps + \frac{3}{8}\eps^2 + \bigO{\eps^3},
\end{align*}
or,
\begin{align*}
\sigma^2 &= 1 - \frac{1}{4s^4} + \frac{3}{8s^8} + \bigO{s^{-12}}
\end{align*}
this approximation is shown in Figure \ref{fig:limsinfty}. \\
\begin{figure}[tbp]
\input{./Figures/Lim_s_Infty}
\caption{Asymptotic expansion of the variance as $s \rightarrow \infty$.}
\label{fig:limsinfty}
\end{figure}

In conclusion, the asymptotic expansions of the observables are
\begin{align*}
\sigma^2 &=
\begin{cases*}
2s(1 - s) + \bigO{s^3} & $s \rightarrow 0$ \\
1 - \frac{1}{4s^4} + \frac{3}{8s^8} + \bigO{s^{-12}} & $s \rightarrow \infty$
\end{cases*}
& \phi &=
\begin{cases*}
\frac{1}{2}s - \frac{1}{24}s^3 + \bigO{s^5} & $s \rightarrow 0$ \\
\frac{\pi}{4} - \frac{1}{2s^2} + \frac{5}{12s^6} + \bigO{s^{-10}} & $s \rightarrow \infty$
\end{cases*} \\
C &=
\begin{cases*}
1 - s + \frac{1}{2}s^2 + \bigO{s^3} & $s \rightarrow 0$ \\
\frac{1}{2s^2} - \frac{3}{8s^6} + \bigO{s^{-10}} & $s \rightarrow \infty$
\end{cases*}
& \zeta &=
\begin{cases*}
1 - \frac{1}{2}s + \frac{1}{8}s^2 + \bigO{s^3} & $s \rightarrow 0$ \\
\frac{1}{4^{1/4}s} - \frac{3 \cdot 4^{3/4}}{32 s^5} + \bigO{s^{-9}} & $s \rightarrow \infty$.
\end{cases*}
\end{align*}
The general form of these is as expected. When $s \rightarrow 0$, we effectively remove the CFBG from the cavity. This causes the pulse to be perpetually modulated, and so, the pulse approaches a $\delta$ function, and therefore, $\sigma^2 \rightarrow 0$. Moreover, since $\zeta$ is the ratio of the amplitude of the pulse before and after dispersion, $\zeta \rightarrow 1$. Similarly, $\phi \rightarrow 0$ since nothing in the circuit has the capacity to induce a phase shift. In the opposite extreme, $s \rightarrow \infty$, we have an incredibly long / powerful CFBG---this causes the pulse to broaden greatly, but still conserving its energy. Because of this, the variance approaches unity because the shape is entirely determined by the modulation function. Additionally, $\zeta \rightarrow 0$ because the pulse becomes infinitely wide, and thus, the amplitude becomes infinitesimally small.

### 2019-05-29 Thesis/Appendices/Hermite.tex
Finished second draft

% !TeX root = ../Thesis.tex
%\chapter{Span of Gaussians in $L^2$}
%\label{chap:gauss}
\section{Span of Gaussians in $L^2$}
In order to show that Gaussians span $L^2$, we shall start our analysis with the span of the Hermite polynomials in $L^2$. Typically the Hermite polynomials are recursively defined as \cite{conway, courant, teuwen}
\begin{align*}
H_n(x) := (-1)^n \textrm{e}^{x^2} \diff[n]{}{x} \textrm{e}^{-x^2},
\end{align*}
with the inner product\footnote{The complex conjugate of $g$ is omitted since the functions dealt with are real.}
\begin{align*}
\left< f, g \right> = \int_\mathbb{R} f(x) g(x) \textrm{e}^{-x^2} \df x,
\end{align*}
where $\textrm{e}^{-x^2}$ is the weighting function. This is so that 
\begin{align*}
\left< H_m, H_n \right> = \sqrt{\pi} 2^n n! \delta_{mn}
\end{align*}
and, therefore, the Hermite polynomials form an orthogonal set \cite{courant, hochstrasser, kreyszig, szego, teuwen}. Consider instead, the Gaussian--Hermite polynomials
\begin{align*}
\Her{n}{x} &:= \textrm{e}^{-x^2/2} H_n(x),
\end{align*}
with the inner product
\begin{align*}
\left< f, g \right> = \int_\mathbb{R} f(x) g(x) \, \df x,
\end{align*}
notice that the weighting function has been absorbed into the Hermite polynomials. We shall now show that the set of Gaussian--Hermite polynomials spans $L^2$. \\

\begin{theorem}
\label{thm:hermite}
The set of Gaussian--Hermite polynomials, $\mathcal{G}(\mathbb{R}) = \{ \Her{n}{x} : x \in \mathbb{R}, n \in \mathbb{N} \}$, spans $L^2(\mathbb{R})$.
\end{theorem}
\begin{proof}
The Gaussian--Hermite polynomials are an orthogonalization of the set $\Xi = \{x^n \textrm{e}^{-x^2/2} : x \in \mathbb{R}, n \in \mathbb{N}\}$, and therefore, have the same span \cite{kreyszig, szego}. Thus, it is sufficient to show that $\Xi$ spans $L^2$.
Suppose there exists an $f \in L^2$ such that $\left< f, \widetilde{H}_n \right> = 0$ for all $n$, or equivalently, that $\left< f, g_n \right> = 0$ for $g_n(x) = x^n \textrm{e}^{-x^2/2}$ for all $n$ so that it is not in the span of $\mathcal{G}(\mathbb{R})$. \\

@@

Alternatively, without much difficulty it can be shown that the Gaussian--Hermite polynomials satisfy
\begin{align}
\label{eq:paracyl}
\diff[2]{\Her{n}{x}}{x} - \left( x^2 - 2n - 1 \right) \Her{n}{x} = 0.
\end{align}
This has the form of the parabolic cylinder functions \cite{miller}, and indeed the Gaussian--Hermite polynomials can be expressed in terms of parabolic cylinder functions \cite{hochstrasser, miller}. Moreover, \ref{eq:paracyl} can also be written in Sturm--Liouville form as
\begin{align*}
\diff{}{x} \left( 1 \cdot \diff{\Her{n}{x}}{x} \right) + (1 - x^2) \Her{n}{x} = -2n \Her{n}{x}.
\end{align*}
By the spectral theorem \cite{griffel, higson, kreyszig}, this suggests that the Gaussian--Hermite polynomials form an orthogonal set with the weighting function $w(x) = 1$, and are complete in $L^2(\mathbb{R})$. Using this result, we can now show that Gaussian functions also span $L^2(\mathbb{R})$ with the following theorem. \\

@@

\begin{figure}[tbp]
\input{./Figures/GaussMod}
\caption[Gaussian series for a particular modulation function.]{$N = 20$, $h = 0.1$}
\label{fig:gaussmod}
\end{figure}

>>>>

% !TeX root = ../Thesis.tex
%\chapter{Span of Gaussians in $L^2$}
%\label{chap:gauss}
\section{Span of Gaussians in $L^2(\mathbb{R})$}
In order to show that Gaussians span $L^2(\mathbb{R})$, we shall start our analysis with the span of the Hermite polynomials in $L^2(\mathbb{R})$. Typically the Hermite polynomials are recursively defined as \cite{conway, courant, teuwen}
\begin{align*}
H_n(x) := (-1)^n \textrm{e}^{x^2} \diff[n]{}{x} \textrm{e}^{-x^2},
\end{align*}
with the inner product\footnote{The complex conjugate of $g$ is omitted since the functions dealt with are real.}
\begin{align*}
\left< f, g \right> = \int_\mathbb{R} f(x) g(x) \textrm{e}^{-x^2} \df x,
\end{align*}
where $\textrm{e}^{-x^2}$ is the weighting function. This is so that 
\begin{align*}
\left< H_m, H_n \right> = \sqrt{\pi} 2^n n! \delta_{mn},
\end{align*}
and therefore, the Hermite polynomials form an orthogonal set \cite{courant, hochstrasser, kreyszig, szego, teuwen}. Consider instead, the Gaussian--Hermite polynomials
\begin{align*}
\Her{n}{x} &:= \textrm{e}^{-x^2/2} H_n(x),
\end{align*}
with the inner product
\begin{align*}
\left< f, g \right> = \int_\mathbb{R} f(x) g(x) \, \df x,
\end{align*}
notice that the weighting function has been absorbed into the Hermite polynomials. We shall now show that the set of Gaussian--Hermite polynomials spans $L^2(\mathbb{R})$. \\

\begin{theorem}
\label{thm:hermite}
The set of Gaussian--Hermite polynomials, $\mathcal{G}(\mathbb{R}) = \{ \Her{n}{x} : x \in \mathbb{R}, n \in \mathbb{N} \}$, spans $L^2(\mathbb{R})$.
\end{theorem}
\begin{proof}
The Gaussian--Hermite polynomials are an orthogonalization of the set $\Xi = \{x^n \textrm{e}^{-x^2/2} : x \in \mathbb{R}, n \in \mathbb{N}\}$, and therefore, have the same span \cite{kreyszig, szego}. Thus, it is sufficient to show that $\Xi$ spans $L^2(\mathbb{R})$.
Suppose there exists an $f \in L^2(\mathbb{R})$ such that $\left< f, \widetilde{H}_n \right> = 0$ for all $n$, or equivalently, that $\left< f, g_n \right> = 0$ for $g_n(x) = x^n \textrm{e}^{-x^2/2}$ for all $n$ so that it is not in the span of $\mathcal{G}(\mathbb{R})$. \\

@@

Alternatively, without much difficulty it can be shown that the Gaussian--Hermite polynomials satisfy
\begin{align}
\label{eq:paracyl}
\diff[2]{\Her{n}{x}}{x} - \left( x^2 - 2n - 1 \right) \Her{n}{x} = 0.
\end{align}
This has the form of the parabolic cylinder functions \cite{miller}, and indeed the Gaussian--Hermite polynomials can be expressed in terms of parabolic cylinder functions \cite{hochstrasser, miller}. Moreover, \eqref{eq:paracyl} can also be written in Sturm--Liouville form as
\begin{align*}
\diff{}{x} \left( 1 \cdot \diff{\Her{n}{x}}{x} \right) + (1 - x^2) \Her{n}{x} = -2n \Her{n}{x}.
\end{align*}
By the spectral theorem \cite{griffel, higson, kreyszig}, this suggests that the Gaussian--Hermite polynomials form an orthogonal set with the weighting function $w(x) = 1$, and are complete in $L^2(\mathbb{R})$. Using this result, we can now show that Gaussian functions also span $L^2(\mathbb{R})$ with the following theorem. \\

@@

\begin{figure}[tbp]
\input{./Figures/GaussMod}
\caption[Gaussian series for a particular modulation function.]{Gaussian series with $N = 20$, $h = 0.1$.}
\label{fig:gaussmod}
\end{figure}
